---
header-includes:
  - \usepackage{wrapfig}
title: "problems in the DHS Covid-19 Flagging methods"
author:
- Marlin Lee, Steve Goldstein, Kyllan Wunder, Abe Megahed
- "University of Wisconsin Data Science Institute - August, 2022"
output:
  pdf_document:
    keep_tex: true
---

Outline:

-   Describe report
-   Describe viz used in analysis
-   flagging noise
-   Trend
-   Day of week


This analysis looks to show three issues in the DHS wastewater and case flagging methods.
These methods have too narrow a scope leading to missing long term trends and over weighting
fluctuations based on noise. Furthermore the case data has a strong day of the week effect 
leading to unjustified flags being created.

```{r set up markdown settings, include =FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	echo = FALSE
)

defOut <- knitr::knit_hooks$get("plot")  # save the default plot hook 
knitr::knit_hooks$set(plot = function(x, options) {  # set new plot hook ...
  x <- defOut(x, options)  # first apply the default hook
  if(!is.null(options$wrapfigure)) {  # then, if option wrapfigure is given ...
    # create the new opening string for the wrapfigure environment ...
    wf <- sprintf("\\begin{wrapfigure}{%s}{%g\\textwidth}", options$wrapfigure[[1]], options$wrapfigure[[2]])
    x  <- gsub("\\begin{figure}", wf, x, fixed = T)  # and replace the default one with it.
    x  <- gsub("{figure}", "{wrapfigure}", x, fixed = T)  # also replace the environment ending
  }
  return(x)
})
```

```{r, out.width= "70%", fig.align = 'right', fig.cap = "Wastewater with flagged dates and quantintle info", fig.align="right", wrapfigure = list("R", .7)}
#, fig.cap = "Wastewater data with flagged days and quantintle info"
library(DSIWastewater)
#load WasteWater_data into the environment
data(WasteWater_data, package = "DSIWastewater")

#restrict Waste data to only Madison data
baseWaste_DF <- WasteWater_data[WasteWater_data$wwtp_name == "Madison MSD WWTF",]

#get DF into format for buildRegressionEstimateTable
baseWaste_DF <-  buildWasteAnalysisDF(baseWaste_DF)

#add quantile data to merge with the regression results
Quantiles_DF <- MakeQuantileColumns(baseWaste_DF,
                                    .8, 90,
                                    "sars_cov2_adj_load_log10")

#Get 5 day rolling regression of data
CDCMethod <- buildRegressionEstimateTable(baseWaste_DF, 
                                          PSigTest=FALSE)
#merge the regression DF and the quantile DF to get info for 

FULL_reg_DF <- dplyr::full_join(Quantiles_DF, CDCMethod,
                                by = c("WWTP", "date"))

#create flags described in @return
FULL_reg_DF <- ClassifyQuantileFlagRegression(FULL_reg_DF)

#return only flags and type columns 
Full_wasteFlags <- FULL_reg_DF[,c("WWTP", "date", 
                                  "sars_cov2_adj_load_log10",
                                  "ntile",
                                  "cdc_flag",
                                  "flag_ntile",
                                  "flag_ntile_pval",
                                  "modeled_percentchange",
                                  "lmreg_sig")]


library(ggplot2)
library(dplyr)

Quantile_Viz_Plot <- Full_wasteFlags%>%
  ggplot(aes(x = date))+
  geom_point(aes(y = 10^(sars_cov2_adj_load_log10)))+
  geom_line(aes(y = 10^(ntile), color = "80th percentile of last 90 days"))+
  geom_point(aes(y = 10^(sars_cov2_adj_load_log10),
                 color = "CDC flag"), 
             data = filter(Full_wasteFlags, cdc_flag == 1))+
  ylab("Covid-19 gene copy per 100k")+
  scale_y_log10()+
  scale_x_date(limits = c(as.Date("2021-01-01"),
                          max(Full_wasteFlags$date)))


Quantile_Viz_Plot
```
# The Data


The department of health services has created flagging methods to to predicatively warn of 
spikes in Wisconsin community's to better inform decision making policy. 

## Wastewater method

The wastewater method uses a two step system to identify potential warning signs in the data. 
The core method is using a five measurement rolling regression to calculate an estimated percent 
change over the time period. Then if it is estimated to be over 100% increase it is logged as a 
CDC flag. Then if the measurement is in the 80th percentile of the last 90 days it is logged as a flag. to the right is a visualization we will be using heavily in this report. It shows the wastewater data
measurements with the color showing the flags with a line showing the quantile info. The flags are the
red points above the line.

```{r, out.width= "65%", fig.align = 'right', fig.cap = "Case slope", fig.align="right", wrapfigure = list("R", .7)}
library(DSIWastewater)

#load Case_data into the environment
data(Case_data, package = "DSIWastewater")


#restrict Case data to only Madison data
Case_DF <- Case_data[Case_data$Site == "Madison",]


#restrict Case data to dates after 2021-02-01
Case_DF <- Case_DF[Case_DF$date  >= as.Date("2021-02-01"),]


#get the case flags
Case_DF <- buildCaseAnalysisDF(Case_DF)


CaseRegressionOutput <- buildRegressionEstimateTable(DataMod = Case_DF, 
    RunOn = c("FirstConfirmed.Per100K", "pastwk.avg.casesperday.Per100K"),
    SplitOn = "Site", DaysRegressed = 7)


library(tidyr)

Case_Flag_Viz_Plot <- CaseRegressionOutput%>%
  select(date, Method, lmreg_slope)%>%
  pivot_wider(names_from = Method, values_from = lmreg_slope)%>%
  ggplot(aes(x = date))+
  geom_line(aes(y = FirstConfirmed.Per100K, color = "cases slope"))+
  geom_line(aes(y = pastwk.avg.casesperday.Per100K, color = "smooth cases slope"))+
  geom_hline(yintercept = 5)+
  ylab("7 Slope")

Case_Flag_Viz_Plot



# both flagging methods have are highly effected by the latent noise in the data. creating flags
# from outliers in the signal. Because the flags are aiming to predict spikes in the community that
# last for many measurements the expected behavior is for flags to happen in sequence. on the right is the wastewater
```

## Case method

The case method uses a 7 day rolling regression to create flags. If the regressed slope is over 5
It is logged as a case flag. Because the slopes are roughly continuous case flags are often repeated
bringing use of the slope switch flag which logged only the first case flag in a sequence. On the right is a visualization we will use in this report. It shows the case slopes vs the seven day smoothing slope data with a horizontal line showing the threshold to be classified as a case flag

# Trend

# Noise


```{r out.width= "65%", fig.align = 'right', fig.cap = "Zoomed in case slope plot", fig.align="right", wrapfigure = list("R", .7)}

Case_Flag_Viz_Plot+
  scale_x_date(limits = c(as.Date("2021-08-10"),
                          as.Date("2021-10-01")))+
  scale_y_continuous(limits = c(-8, 10))+
  geom_vline(aes(xintercept = date, color = "Friday"),
             data = filter(CaseRegressionOutput, weekdays(date) == "Friday"))
```

# Day of week effect

The issue is subtle in how it effects the flagging method but close examination clearly shows a strong effect. in this time period we show a strong effect with the flag on September 17 being partial or
entirely generated by this effect.