---
header-includes:
- \usepackage{wrapfig}
title: "day of week effect in DHS COVID-19 case Flagging Method"
author:
- Steve Goldstein, Marlin Lee, Abe Megahed, Kyllan Wunder
- "University of Wisconsin Data Science Institute - August, 2022"
output:
  pdf_document:
    keep_tex: true     
editor_options: 
  chunk_output_type: console
---

```{r set up markdown settings, echo=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	echo = FALSE
)
```


```{r}
library(Covid19Wastewater)

#load Case_data into the environment
data(Case_data, package = "Covid19Wastewater")


#get the case flags
Case_DF <- buildCaseAnalysisDF(Case_data)


CaseRegressionOutput <- buildRegressionEstimateTable(DataMod = Case_DF, 
    RunOn = c("FirstConfirmed.Per100K", "pastwk.avg.casesperday.Per100K"),
    SplitOn = "site", DaysRegressed = 7, verbose = FALSE)

```


```{r}
library(lubridate)
library(dplyr)
library(tidyr)

DOWCaseRegression <- CaseRegressionOutput%>%
  mutate(NumWeek = week(date), dayweek = wday(date, label = TRUE))
```

This is an extension of the analysis shown on 9/8. We aim to show that the case flag method
has a meaningful day of the week effect across the whole data set.

The first table shows the mean slope of each day of the week. with Friday having a mean
slope of 2 and Sunday having mean slope of -1.7 there is a huge difference between days


```{r}
DOWCaseRegression%>%
  group_by(Method, dayweek)%>%
  summarise(meanSlope = mean(lmreg_slope))%>%
  pivot_wider(names_from  = Method, values_from  = c(meanSlope))

#show over flaging
```
The second table has the percent each weekday as a flag. having a 3x difference is large
enough to have concern that this could be affecting the flagging method

```{r}
DOWCaseRegression%>%
  mutate(flag = lmreg_slope>5)%>%
  group_by(Method, dayweek)%>%
  summarise(meanflag = 100*mean(flag))%>%
  pivot_wider(names_from  = Method, values_from  = meanflag)
```
