---
title: "HFGCrossTrend"
author: "Marlin"
date: "2022-10-21"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Paper:
using HFD to get a good understanding of the relationships between the Covid-19
measurements and the other covariates. we did this by breaking the signal down into
three components. (1) the trend of the data. this is created using a loess smoothing.
(2) the first difference of the trend. (3) the detrended data. this captures the noise
processes. furthermore there is two view of the data. We look at the normal data
and the log of data. The log of the data better captures the shape of Wastewater.
at the bottom of the document contain plots showing a comprehensive view of the
problem. The main goal of this document is to show the main takeaways of those plots
in an easier way to see. The main takeaways are, 1) the covariates don't strongly
relate to the covid signal. 2) Crp and HF183 are strongly related. 3)

```{r cars}
library(DSIWastewater)
library(dplyr)
library(ggplot2)
library(tidyr)
data("HFGWaste_data", package = "DSIWastewater")

```

```{r}

#crazy agressive method
hfg_outlier_detection <- function(small_vec){
  upbound_vec <- quantile(small_vec, probs = .7, na.rm = TRUE)
  retVec = ifelse(small_vec > upbound_vec,
                  NA, small_vec)
  retVec = ifelse(retVec == 0, NA, retVec)
  return(retVec)
}

Pop_DF <- data.frame(
  site = c("Hudson","Kenosha","Platteville","Madison","Merrill","Plymouth","River Falls","Sun Prairie","Wausau","Oshkosh","Wausau"),
  pop = c(19680,122000,14000,380000,10000,9000,16000,34926,42000,67000,42000)
)

hfg_waste_filt_df <- HFGWaste_data%>%
  select(site, date, Filter, Well, N1, N2, PMMOV, BCoV, HF183, CrP)%>%
  group_by(date,site)%>%
  mutate(across(N1:CrP, hfg_outlier_detection))%>%
  left_join(Pop_DF)
```

```{r}
trend_df <- hfg_waste_filt_df%>%
  group_by(site)%>%
  group_split()%>%
  lapply(loessSmoothMod, InVar = "N1", OutVar = "N1_Trend")%>%
  lapply(loessSmoothMod, InVar = "N2", OutVar = "N2_Trend")%>%
  lapply(loessSmoothMod, InVar = "PMMOV", OutVar = "PMMOV_Trend")%>%
  lapply(loessSmoothMod, InVar = "HF183", OutVar = "HF183_Trend")%>%
  lapply(loessSmoothMod, InVar = "CrP", OutVar = "CrP_Trend")%>%
  bind_rows()%>%
  mutate(
    N1_Diff = N1_Trend - N1,
    N2_Diff = N2_Trend - N2,
    PMMOV_Diff = PMMOV_Trend - PMMOV,
    HF183_Diff = HF183_Trend - HF183,
    CrP_Diff = CrP_Trend - CrP
    )%>%
  select(date, site, pop, N1_Trend:CrP_Diff)
```



```{r}
log_trend_df <- hfg_waste_filt_df%>%
  mutate(log_N1 = log(N1),
         log_N2 = log(N2),
         log_PMMOV = log(PMMOV),
         log_HF183 = log(HF183),
         log_CrP = log(CrP))%>%
  group_by(site)%>%
  group_split()%>%
  lapply(loessSmoothMod, InVar = "log_N1", OutVar = "log_N1_Trend")%>%
  lapply(loessSmoothMod, InVar = "log_N2", OutVar = "log_N2_Trend")%>%
  lapply(loessSmoothMod, InVar = "log_PMMOV", OutVar = "log_PMMOV_Trend")%>%
  lapply(loessSmoothMod, InVar = "log_HF183", OutVar = "log_HF183_Trend")%>%
  lapply(loessSmoothMod, InVar = "log_CrP", OutVar = "log_CrP_Trend")%>%
  bind_rows()%>%
  mutate(
    N1_Diff = log_N1_Trend - log_N1,
    N2_Diff = log_N2_Trend - log_N2,
    PMMOV_Diff = log_PMMOV_Trend - log_PMMOV,
    HF183_Diff = log_HF183_Trend - log_HF183,
    CrP_Diff = log_CrP_Trend - log_CrP)%>%
  select(site, date, pop, log_N1_Trend:CrP_Diff)
```
```{r}
trend_df
  
  

log_trend_df
```


```{r fig.height = 15, fig.width= 15}
library(ggcorrplot)
library(GGally)

trend_df%>%
  mutate(date = as.numeric(date))%>%
  select(date, pop:CrP_Diff)%>%
  cor(use = "complete.obs")%>%
  ggcorrplot()
trend_df%>%
  ggpairs(aes(colour = site, alpha = 0.4))


log_trend_df%>%
  select(date, pop, log_N1_Trend:CrP_Diff)%>%
  mutate(date = as.numeric(date))%>%
  cor(use = "complete.obs")%>%
  ggcorrplot()
log_trend_df%>%
  ggpairs(aes(colour = site, alpha = 0.4))
```