---
title: "Heuristic exploration of the relationship between FirstConfirmed and viral load"
author: "Marlin Lee"
date: "1/28/2022"
output:
  pdf_document: null
  bookdown::html_document2:
    fig_width: 8
    fig_height: 4
  html_document:
    df_print: paged
always_allow_html: yes
editor_options:
  chunk_output_type: inline
---

**This report looks is an update to the analysis shown on 1/14/2022. Most steps are the same with tweaking due to different data and outlier procedure**

at exploring the relationship between wastewater and FirstConfirmed. 
There are four components to this analysis.

  1) Removing putative outliers
  
  2) Binning analysis
  
  3) Smoothing signal
  
  4) Statistical analysis

  This report does not present any final answers but presents some very convincing heuristics.



```{r set up markdown settings, echo=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	echo = FALSE
)

output.format <- knitr::opts_knit$get("rmarkdown.pandoc.to")
```


```{r Start enviroment, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#Used library for visualization
library(dplyr)
library(ggplot2)
library(lmtest)
library(lubridate)
#library(limma)
library(zoo)

if ("html" != output.format) {
  
}else{
  library(plotly)
  library(formattable)
}
```



```{r helpful data manipulations, include = FALSE}

#Custom_color_scale
PlotColors <- c("#F8766D", "#00BFC4", "#4057A2", "#999999", "#D6544B")
PlotObjects <- c("Wastewater","FirstConfirmed","Seven Day MA FirstConfirmed",
                 "Wastewater outlier","loess Wastewater")

ColorRule <- scale_color_manual(
  values = setNames(as.list(PlotColors),PlotObjects))
```

"data Used from DSIWastewater package"


```{r DF Set Up, echo=FALSE}
library(DSIWastewater)
data(Case_data, package = "DSIWastewater")

#restrict Case data to only Madison data
Case_DF <- Case_data[Case_data$Site == "Madison",]


#restrict Case data to dates after 2021-02-01
Case_DF <- Case_DF[Case_DF$date  >= as.Date("2020-9-10"),]

#get the case flags
Case_DF <- buildCaseAnalysisDF(Case_DF)

Case_DF <- Case_DF[,c(1:2, 5, 7)]

data(WasteWater_data, package = "DSIWastewater")

#restrict Waste data to only Madison data
baseWaste_DF <- WasteWater_data[WasteWater_data$wwtp_name == "Madison MSD WWTF",]

#get DF into format for buildRegressionEstimateTable
baseWaste_DF <-  buildWasteAnalysisDF(baseWaste_DF)

#Select columns needed in analysis

LIMSFullDF <- baseWaste_DF[,c(1:2,7)]

#joining the two data frames together
FullDF <- full_join(LatCaseDF,LIMSFullDF, by = c("Date","Site"))
```

# Data: The first look

The two data sets used in this analysis are the Madison case data sourced from the Wisconsin DHS and wastewater  concentration data produced by the Wisconsin State Laboratory of Hygiene. This wastewater data has entries every couple of days from `r format(min(LIMSFullDF$Date), "%d %B %Y")` to `r format(max(LIMSFullDF$Date), "%d %B %Y")`.

```{r First look, echo=FALSE}
#what does the data look like? 
Disp1 <- FullDF%>%
  filter(!is.na(sars_cov2_adj_load))%>%
  head()

#diff()
AVGLimsSample <- LIMSFullDF%>%
  filter(!is.na(sars_cov2_adj_load))%>%
  pull(Date)%>%
  sort()%>%
  diff()%>%
  mean()%>%
  as.numeric()

FullDF%>%
  filter(!is.na(sars_cov2_adj_load))%>%
  summarise(min(Date),max(Date))
  
if ("html" != output.format) {
  Disp1%>%
    knitr::kable()
}else{
  Disp1%>%
    formattable()
}

```

The case data has a strong weekend effect so for this section we look at a seven day smoothing of FirstConfirmed. The simple display of the data shows the core components of this story. First, wastewater data is noisy. And that there is a clear relationship between the two signals.

```{r fig.cap= 'Wastewater concentration and daily Covid-19 case data for Madison. A seven day moving average of FirstConfirmed is used to reduce a day of the week effect.'}
FirstImpressionDF <- FullDF%>%
  filter(!is.na(sars_cov2_adj_load),
         !is.na(FirstConfirmed))#Removing NA

FirstImpression <- FirstImpressionDF%>%
  ggplot(aes(x = Date))+#Data depends on time
  geom_point(aes(y = FirstConfirmed, color = "FirstConfirmed", info = FirstConfirmed),
             size = 1)+
  geom_point(aes(y = sars_cov2_adj_load, 
                color = "Wastewater",
                info = sars_cov2_adj_load),
             size = 1)+#compares sars_cov2_adj_load to FirstConfirmed
  geom_line(aes( y= MAWaste, 
                color = "Seven Day MA Waste",
                info = FirstConfirmed))+
  geom_line(aes(y = MAFirstConfirmed, 
                color = "Seven Day MA FirstConfirmed",
                info = FirstConfirmed))+
  labs(y = "Reported FirstConfirmed")+
  scale_y_log10()


if ("html" != output.format) {
  PrintPlot <- FirstImpression
}else{
  PrintPlot <- ggplotly(FirstImpression)
}
PrintPlot
#To remove weekend effects we are looking at the 7 day smoothing of FirstConfirmed.
```
  

# Removing potential outliers

Looking at the wastewater measurements we observe there were some points many times larger than adjacent values hinting at them being outliers. We used the adjacent 10 values on each side and marked points 2.5 standard deviations away from the group mean as outliers.

```{r fig.cap= 'Wastewater concentration for Madison with potential outliers marked. Using a rolling symmetrical bin of 21 days as a sample we use 2.5 standard deviations of the bin as a metric to reject extreme points. This process is ran multiple times to get a robust process to select outliers.'}
#default pass to IdentifyOutliers
#method="SD", align="center", n = 5, Bin = 21, Action = "Flag"

source("OutlierDetectionFuncs.R")

ErrorMarkedDF <- FullDF%>%#
    mutate(FlagedOutliers = IdentifyOutliers(sars_cov2_adj_load, Action = "Flag"),
           #Manual flagging that method misses due to boundary effect with binning
           NoOutlierVar = ifelse(FlagedOutliers, NA, sars_cov2_adj_load))

#Split N1 into outlier and non outlier for next ggplot
OutLierPlotDF <- ErrorMarkedDF%>%#
  mutate(outlier_sars_cov2_adj_load = ifelse(FlagedOutliers, sars_cov2_adj_load,NA))%>%
           mutate(sars_cov2_adj_load = NoOutlierVar)%>%
  arrange(Date)%>%
  mutate(MAWasteWOOutlier = rollmean(sars_cov2_adj_load, 7, align = "right",
                            na.rm = TRUE,fill=NA))

#Drop Var create Var filter 
UpdatedDF <- ErrorMarkedDF%>%
  select(-sars_cov2_adj_load)%>%
  rename(sars_cov2_adj_load = NoOutlierVar)%>%
  arrange(Date)%>%
  mutate(MAWaste = rollmean(sars_cov2_adj_load, 7, align = "right",
                            na.rm = TRUE,fill=NA))

OutLierPlotObject <- OutLierPlotDF%>%
  filter(!(is.na(sars_cov2_adj_load)&is.na(outlier_sars_cov2_adj_load)))%>%
  ggplot(aes(x=Date))+#Data depends on time 
  geom_point(aes(y=sars_cov2_adj_load,
                color="sars_cov2_adj_load", 
                info = sars_cov2_adj_load),
             size = 1)+#compares Var to FirstConfirmed
  geom_line(aes(y= MAWaste, 
                color="Seven Day MA Waste",
                info=FirstConfirmed))+
  geom_line(aes(y= MAWasteWOOutlier, 
                color="Seven Day MA Waste W/O Outlier",
                info=FirstConfirmed))+
  geom_point(aes(y = outlier_sars_cov2_adj_load,
                 color= "outlier_sars_cov2_adj_load",
                 info = outlier_sars_cov2_adj_load))+
  scale_y_log10()

if ("html" != output.format) {
  PrintPlot <- OutLierPlotObject
}else{
  PrintPlot <- ggplotly(OutLierPlotObject, tooltip=c("info","Date"))
}
PrintPlot
```


# Data smoothing

The goal in this section is to smooth the data to get a similar effect without losing
resolution.


```{r, include=FALSE}

SpanConstant = .06

AVGLimsSample*length(LIMSFullDF$sars_cov2_adj_load)*SpanConstant/14
```



## viral load smoothing

To get a good smoothing of the sars_cov2_adj_load measurement we employ loess smoothing. Loess smoothing takes a locally weighted sliding window using some number of points. we found the best smoothing when it uses data within approximately `r floor(AVGLimsSample*length(LIMSFullDF$N1)*SpanConstant/14)` weeks of both sides of the data. The displayed plot shows the visual power of this smoothing. We see in general that the smoothed N1 trails SLD. However loess is symmetric meaning that it can not be used in predictive modeling due to it using points from the future to smooth points.



```{r fig.cap= 'Loess smoothed N1 and SLD FirstConfirmed for Madison data. Using a Locally Weighted Scatterplot Smoothing process along with the previous figure SLD FirstConfirmed we get the most sophisticated relationship between the two signals discussed in this document.'}
SpanConstant = .06

UpdatedDFLoess <- LoessSmoothMod(UpdatedDF,
                                "sars_cov2_adj_load",
                                "loess_sars_cov2_adj_load",
                                span = SpanConstant)


LoessGraphic <- UpdatedDF%>%
  filter(!is.na(loess_sars_cov2_adj_load))%>%
  ggplot(aes(x = Date))+
  geom_point(aes(y = FirstConfirmed, color = "FirstConfirmed" , info = FirstConfirmed),
             alpha = .1,
             size = 1)+
  geom_point(aes(y = sars_cov2_adj_load,
                color = "sars_cov2_adj_load",
                info = sars_cov2_adj_load),
            alpha = .2)+
  geom_line(aes(y = MAWaste, 
                color = "Seven Day MA Waste",
                info = FirstConfirmed),
            alpha = .3)+
  geom_line(aes(y = MAFirstConfirmed,
                color = "Seven Day MA FirstConfirmed" , 
                info = MAFirstConfirmed))+
  geom_line(aes(y = loess_sars_cov2_adj_load, 
                color = "loess_sars_cov2_adj_load" ,
                info = loess_sars_cov2_adj_load))+
  labs(y = "Reported FirstConfirmed")+
  scale_y_log10()#+
  #ColorRule



if ("html" != output.format) {
  PrintPlot <- LoessGraphic
}else{
  PrintPlot <- ggplotly(LoessGraphic,tooltip=c("info","Date"))
}
PrintPlot
```


# Towards a formal analysis


Cross correlation and Granger Causality are key components to formalize this analysis. Cross correlation looks at the correlation at a range of time shifts and Granger analysis performs a test for predictive power.


```{r}
CCFChar <- function(ccfObject){
  LargestC = max(ccfObject$acf)
  
  Lag = which.max(ccfObject$acf) - 21
  
  return(c(LargestC,Lag))
}

ModelTesting <- function(DF,Var1,Var2){
  #removing rows from before both series started
  UsedDF <- DF%>%
    filter(!is.na(Var1),
           !is.na(Var2))
  
  
  Vec1 <- log(unname(unlist(UsedDF[Var1])))
  
  Vec2 <- log(unname(unlist(UsedDF[Var2])))

  CCFReport <- CCFChar(ccf(Vec1,Vec2,na.action=na.pass,plot = FALSE))
  
  ShiftSize = 3#CCFReport[2]
  
  LMDF <- data.frame(
    Ver1 = Vec1,
    Ver2 = Vec2
  )
  
  VarPredCase <- grangertest(formula = Ver2 ~ Ver1,
                             order = ShiftSize,
                             data = LMDF)$"Pr(>F)"[2]

  CasePredVar <- grangertest(formula = Ver1 ~ Ver2,
                             order = ShiftSize,
                             data = LMDF)$"Pr(>F)"[2]
  

  return(round(c(CCFReport, VarPredCase, CasePredVar),4))
}

#ErrorRemovedDF
BaseLine <- ModelTesting(FullDF, "sars_cov2_adj_load", "FirstConfirmed")
BaseLineSevenDay <- ModelTesting(FullDF, "sars_cov2_adj_load", "MAFirstConfirmed")
ErrorRemoved <- ModelTesting(UpdatedDF, "sars_cov2_adj_load", "MAFirstConfirmed")
SevenLoess <- ModelTesting(UpdatedDF, "loess_sars_cov2_adj_load","MAFirstConfirmed")


Output <- data.frame(row.names=c("Max Cross Correlation",
                                 "Lag of largest Cross correlation",
                                 "P-value Wastewater predicts FirstConfirmed",
                                 "P-value FirstConfirmed predicts wastewater"),
  FirstConfirmedvsWaste = BaseLine,
  MAFirstConfirmedvsWaste = BaseLineSevenDay,
  MAFirstConfirmedvsErrorRemovedWaste = ErrorRemoved,
  SevenLoess = SevenLoess)

OutputRightPosition <- data.frame(t(Output))

colnames(OutputRightPosition) <- rownames(Output)

rownames(OutputRightPosition) <- c(paste("Section 1: FirstConfirmed vs sars_cov2_adj_load"),
                                   paste("Section 1: 7 Day MA FirstConfirmed vs sars_cov2_adj_load"),
                                  paste("Section 2: FirstConfirmed vs sars_cov2_adj_load"),
                                   paste("Section 4.3: 7 Day MA FirstConfirmed vs Loess smoothing of sars_cov2_adj_load"))

if ("html" != output.format) {
  OutputRightPosition%>%
    knitr::kable()
}else{
  OutputRightPosition%>%
    formattable()
}
```

